Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web

### 2024-11-19 autovi_paper.tex
done with cutting paper

\subsection{Numerical Summary}\label{sec-autovi-usage}

@@

\subsection{Visual Summary}\label{sec-autovi-visual}

@@

The \texttt{autovi.web} package extends the functionality of
\texttt{autovi} by offering a user-friendly web interface for automated
residual plot assessment. This eliminates the common challenges
associated with software installation, so users can avoid managing
Python environments or handling version requirements for R libraries.
The platform is cross-platform and accessible on various devices and
operating systems, making it suitable even for users without R
programming experience. Additionally, updates are managed centrally,
ensuring that users always have access to the latest features.

The \texttt{autovi.web} interface is available at
\url{autoviweb.netlify.app}. This section discusses the implementation
based on \texttt{autovi.web} version 0.1.0.

@@

\subsection{Design}\label{sec-autovi-web-design}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/autovi_web.png}

@@

\caption{\label{fig-autovi-web}Overview of the \texttt{autovi.web}
graphical user interface (GUI). This default view may change based on
user interactions. Region 1 is the sidebar menu, containing the residual
assessment tab and the information tab. Region 2 is the data upload
panel, where users can provide a CSV file and specify the type of data
it contains. Region 3 includes dropdown menus for selecting the columns
to be analyzed, a slider to control the number of bootstrapping samples,
and a numeric input box for setting the simulation seed. Region 4
displays the initialization status and offers a button to start the
analysis. Region 5 is empty in the default view but will be populated
with results once the analysis is started.}

@@

While the R package \texttt{autovi} aims to provide tools that can be
extended to broader visual inference applications, \texttt{autovi.web}
is only focus on providing a straightforward and clean user interface.
An overview of the graphical user interface of \texttt{autovi.web} is
provided in Figure~\ref{fig-autovi-web}. This is the default view of the
web application, and there are five regions that user can mainly
interact with. Region 1 of Figure~\ref{fig-autovi-web} is a sidebar menu
which can switch between the analysis page and the information page. The
analysis page is the focus of this section.

Region 2 of Figure~\ref{fig-autovi-web} is a panel for data uploading
and CSV type selection. Clicking the ``upload CSV'' button opens a
window where the user can select a file from their local system. The
data status displayed above the button provides information about the
number of rows and columns in the current dataset. Additionally, there
are two example datasets available beneath the ``upload CSV'' button:
one is a lineup example using a CSV file with three columns, and the
other is a single plot example using a CSV file with two columns. More
details about these example datasets are be discussed in
Section~\ref{sec-autovi-web-workflow}.

While the \texttt{autovi} package typically expects a fitted regression
model object provided by the user, this approach is impractical for a
web interface. Saving the R model object to the filesystem involves
extra steps and requires users to have specific knowledge, which does
not align with the goal of the web application. Moreover, the regression
model object may contain sensitive, non-shareable data, making it
unsuitable for uploading. Additionally, model objects are often
unnecessarily large, containing extra information not needed for
residual diagnostics. In contrast, a CSV file is easier to generate
using various software programs, not just R. CSV files are widely
accepted and can be easily viewed and modified using common desktop
applications like Excel. They are generally less sensitive than raw
data, as they exclude most information about the predictors.

The web application is designed to assess either a single residual plot
or a lineup of residual plots. Therefore, it accepts only two types of
CSV files: one with at least two columns representing the fitted values
and residuals of a single residual plot, and another with at least three
columns, where the additional column serves as the label or identifier
for a lineup of multiple residual plots. For a single residual plot, 19
null plots are generated by simulating normal random draws from a
distribution with the same variance as the original residual plot, and
comparisons are made with the original residual plot. For a lineup,
comparisons are made among the plots within the lineup. After uploading
the CSV file, the user must select the correct format to ensure the web
interface interprets the data correctly.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/autovi_web_type.png}

@@

\caption{\label{fig-autovi-web-type}The panels for selecting target
columns and simulation settings are updated when a different CSV type is
selected in the left panel. Compared to Figure~\ref{fig-autovi-web},
where the CSV type is a single residual plot, choosing a CSV type that
includes a lineup of multiple residual plots adds a dropdown menu for
specifying a column for the residual plot identifier. Additionally, an
optional dropdown menu for specifying the true residual plot identifier
will appear under the simulation settings.}

Region 3 of Figure~\ref{fig-autovi-web} is a panel for column selection
and simulation settings. As shown in Figure~\ref{fig-autovi-web}, if the
CSV type is set to a single residual plot, there will be two dropdown
menus for specifying the columns for fitted values and residuals,
respectively. The default variable names for these columns are
\texttt{.fitted} and \texttt{.resid}. After uploading the CSV file, the
content of these dropdown menus will be updated to reflect the existing
columns in the dataset. As displayed in
Figure~\ref{fig-autovi-web-type}, for the CSV type that is a lineup of
multiple residual plots, an additional dropdown menu will appear for
specifying the column of residual plot labels. The default variable name
for this column is \texttt{.sample}. If this variable name does not
exist in the dataset, the dropdown menu will remain empty, allowing the
user to specify the correct column. The number of levels for each option
in this dropdown menu will be displayed to help avoid the selection of a
variable with too many levels, which could significantly slow down the
application due to extensive computation.

Under the simulation settings, there is a slider for specifying the
number of bootstrapped samples needed for the assessment. A higher value
on this slider will result in a more accurate bootstrap distribution
estimation, though it will require more computation time. The simulation
seed can be set in a numeric input box below the slider to control the
reproducibility of the assessment. By default, a random seed is set each
time the web page is refreshed. When the CSV type is a lineup of
multiple residual plots, an optional dropdown menu will appear next to
the simulation seed input box, allowing the user to specify an
identifier for the true residual plot. If no label is provided for the
true residual plot, the assessment will only estimate the VSS for each
residual plot in the lineup, without providing a \(p\)-value, as it
cannot be computed. Consequently, some result panels may be missing due
to insufficient information. This option is useful when the lineup
consists solely of null plots or if the user simply wants to obtain the
VSS for multiple residual plots.

Region 4 of Figure~\ref{fig-autovi-web} is the panel for triggering the
assessment. It contains a large play button to start the assessment.
Above the play button, a text message displays the status of
\texttt{TensorFlow.js}, allowing users to monitor whether the JavaScript
library and Keras model have been loaded correctly. The play button will
remain disabled until both the data status in Region 1 and the
\texttt{TensorFlow.js} status in Region 4 indicate that everything is
ready, with both showing a green status.

Once the play button is clicked, region 5 of Figure~\ref{fig-autovi-web}
will be populated with panels displaying the assessment results.
Generally, there will be four result panels, as shown in
Figure~\ref{fig-autovi-web-result} and
Figure~\ref{fig-autovi-web-result2}.

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/autovi_web_result.png}

\caption{\label{fig-autovi-web-result}The first two panels of results
from the automated residual assessment are shown. The application
provides four results panels in total, and these screenshots display the
first two. In region 1, there is an interactive table detailing the VSS,
with a summary of the analysis provided in the paragraph below. Region 2
displays a lineup of residual plots.}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/autovi_web_result2.png}

@@

\caption{\label{fig-autovi-web-result2}The last two panels of results
from the automated residual assessment are shown. The application
provides four results panels in total, and these screenshots display the
final two. Region 1 presents a density plot comparing the bootstrapped
VSS with the null VSS. Region 2 includes an attention map of the true
residual plot.}

\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{figures/autovi_web_gradient_hide.png}

\caption{\label{fig-autovi-web-gradient-hide}The attention map is hidden
if the assessment indicates a \(p\)-value greater than 0.05. A button is
available to toggle the display of the attention map.}

Region 6 of Figure~\ref{fig-autovi-web-result} contains an interactive
table created with the R package \texttt{DT} \citep{dt}, which provides
the VSS. This table includes four columns: \texttt{.sample},
\texttt{vss}, \texttt{rank}, and \texttt{null}. The \texttt{.sample}
column shows the residual plot labels. For a CSV type that is a lineup,
these labels are taken from an identifier column in the dataset
specified by the user. In the case of the CSV type is a single residual
plot, labels are automatically generated from 1 to 20, with the true
residual plot receiving a randomly assigned label. The \texttt{vss}
column displays the VSS for each residual plot, rounded to three decimal
places. The \texttt{rank} column indicates the ranking of each residual
plot based on VSS. The \texttt{null} column reveals whether the plot is
a null plot. For the CSV type that is a single residual plot, only the
true residual plot will have ``false'' in this column, while all other
plots will be marked ``true.'' For the CSV type that is a lineup, if the
true residual plot identifier has not been provided, this column will
show ``NA'' to represent missing values. If the identifier is provided
by user, the column behaves as if the CSV type is a single residual
plot.

The \texttt{DT} table provides several interactive features. Users can
download the table in four formats, including text, CSV, Excel, and PDF,
using the buttons located above the table. Additionally, the table is
searchable via the text input field also positioned above it. Below the
table, a text message displays the \(p\)-value of the assessment for the
true residual plot and summarizes the number of null plots with VSS
greater than that of the true residual plot. This helps the user
determine whether the true residual plot shows visual patterns that
suggest model violations.

Region 7 of Figure~\ref{fig-autovi-web-result} provides a lineup of
plots corresponding to each \texttt{.sample} value from the table in
Region 6. Due to space limitations, a maximum of 20 residual plots will
be displayed, ensuring that the true residual plot, if known, will be
included in the lineup. The plots are generated using \texttt{ggplot2},
the same as in \texttt{autovi}. Users can perform a visual test with
this lineup to check if the true residual plot is distinguishable from
the other plots, helping to determine the significance of model
violations.

Region 8 of Figure~\ref{fig-autovi-web-result2} displays the density
plot for bootstrapped VSS and null VSS. The densities are shown in
distinct colors that are friendly for colorblind users. A solid vertical
line marks the VSS of the true residual plot, while rug lines at the
bottom of the plot provide a clearer view of individual cases. Below the
plot, a text message indicates the number and percentage of bootstrapped
residual plots that would be rejected by the visual test when compared
to the null plots. Note that the bootstrapped residual plots in this
application are generated differently from \texttt{autovi}. Since we do
not have the R model object, we can not refit the regression model with
bootstrapped data. Instead, we bootstrap the residuals of the true
residual plot directly to obtain bootstrapped residual plots. As as
result, this panel will disappear when the true residual plot is
unknown.

Region 9 of Figure~\ref{fig-autovi-web-result2} displays an attention
map for the true residual plot, generated by computing the gradient of
the Keras model's output with respect to the greyscale input of the
plot. The attention map helps to understand how the Keras model predicts
VSS and which areas it is focusing on. We use a greyscale input because
it is easier to generate a clear attention map in this format, and it
usually conveys all the essential information, as most of the important
details of the plot are drawn in black. If the \(p\)-value of the true
residual plot is greater than 0.05, checking the attention map is not
necessary. However, to provide users with the option to review it if
they wish, a button will be available, as shown in
Figure~\ref{fig-autovi-web-gradient-hide}. This button allows users to
toggle the display of the attention map.

\subsection{Workflow}\label{sec-autovi-web-workflow}

The workflow of \texttt{autovi.web} is designed to be straightforward,
with numbered steps displayed in each panel shown in
Figure~\ref{fig-autovi-web}. There are two example datasets provided by
the web application, as mentioned in
Section~\ref{sec-autovi-web-design}. The single residual plot example
uses the \texttt{dino} dataset from the R package \texttt{datasauRus}
\citep{datasaurus}. The lineup example uses residuals from a simulated
regression model that has a non-linearity issue. We will walk through
the lineup example to further demonstrate the workflow of the web
application.

As shown in Figure~\ref{fig-autovi-web-workflow-example1}, to use the
lineup example data, click the ``Use Lineup Example'' button. The data
status will then update to show the number of rows and columns in the
dataset, and the CSV type will automatically be selected to the correct
option. Since the example dataset follows the variable naming
conventions assumed by the web application, the columns for fitted
values, residuals, and labels of residual plots are set automatically
(Figure~\ref{fig-autovi-web-workflow-example2}). If the user is working
with a custom dataset, these options must be set accordingly. Regardless
of the dataset, the user must manually select the label for the true
residual plot, as the web application allows assessments without this
label. The next step is to click the play button to start the
assessment.

Results are provided in multiple panels as displayed in
Figure~\ref{fig-autovi-web-workflow-example3},
Figure~\ref{fig-autovi-web-workflow-example4},
Figure~\ref{fig-autovi-web-workflow-example5} and
Figure~\ref{fig-autovi-web-workflow-example6}. In
Figure~\ref{fig-autovi-web-workflow-example3}, the first row of the
table is the most crucial to check, as it provides the VSS and the rank
of the true residual plot among the other plots. The summary text
beneath the table provides the \(p\)-value, which can be used for quick
decision-making. In Figure~\ref{fig-autovi-web-workflow-example4}, the
lineup is for manual inspection, and the user should see if the true
residual plot is visually distinguishable from the other plots, to
confirm if the model violation is serious. The density plot in
Figure~\ref{fig-autovi-web-workflow-example5} offers a more robust
result, allowing the user to compare the distribution of bootstrapped
VSS with the distribution of null VSS. Finally, the grayscale attention
map shown in Figure~\ref{fig-autovi-web-workflow-example6} can be used
to check if the target visual features, like the non-linearity present
in the lineup example, are captured by the computer vision model,
ensuring the quality of the assessment.

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-workflow-example1-1.png}

\caption{\label{fig-autovi-web-workflow-example1}To begin the workflow
for \texttt{autovi} using the lineup example dataset, the user clicks
the ``Use Lineup Example'' button to load the example dataset, during
which the data status and CSV type will be automatically updated.}

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-workflow-example2-1.png}

\caption{\label{fig-autovi-web-workflow-example2}After clicking the
button in Figure~\ref{fig-autovi-web-workflow-example1}, the target
columns are selected automatically, though the user must manually select
the label for the true residual plot, as the web application permits
assessment without this label.}

\includegraphics[width=0.55\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-workflow-example3-1.png}

\caption{\label{fig-autovi-web-workflow-example3}After finishing the
required steps in Figure~\ref{fig-autovi-web-workflow-example1} and
Figure~\ref{fig-autovi-web-workflow-example2}, the user initiates the
assessment of the lineup example data by clicking the run button.}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-workflow-example4-1.png}

\caption{\label{fig-autovi-web-workflow-example4}The VSS of the true
residual plot is displayed in the first row of the table, with a summary
text beneath the table providing the \(p\)-value to aid in
decision-making.}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-workflow-example5-1.png}

\caption{\label{fig-autovi-web-workflow-example5}A lineup of residual
plots allows for manual inspection.}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-workflow-example6-1.png}

\caption{\label{fig-autovi-web-workflow-example6}The density plot helps
verify if the bootstrapped distribution differs from the null
distribution.}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-workflow-example7-1.png}

\caption{\label{fig-autovi-web-workflow-example7}The attention map
offers insights into whether the computer vision model has captured the
intended visual features of the true residual plot.}

>>>>

\subsection{Usage}\label{sec-autovi-usage}

\subsubsection{Numerical Summary}\label{sec-autovi-numerical}

@@

\subsubsection{Visual Summary}\label{sec-autovi-visual}

@@

The \texttt{autovi.web} package extends the functionality of
\texttt{autovi} by offering a user-friendly web interface for automated
residual plot assessment. This eliminates the common challenges
associated with software installation, so users can avoid managing
Python environments or handling version requirements for R libraries.
The platform is cross-platform and accessible on various devices and
operating systems, making it suitable even for users without R
programming experience. Additionally, updates are managed centrally,
ensuring that users always have access to the latest features. This
section discusses the implementation based on \texttt{autovi.web}
version 0.1.0.

@@

\subsection{Usage}\label{sec-autovi-web-workflow}

The workflow of \texttt{autovi.web} is designed to be straightforward,
with numbered steps displayed in each panel. There are two example
datasets provided by the web application. The single residual plot
example uses the \texttt{dino} dataset from the R package
\texttt{datasauRus} \citep{datasaurus}. The lineup example uses
residuals from a simulated regression model that has a non-linearity
issue. We will walk through the lineup example to further demonstrate
the workflow of the web application.

\subsubsection{Reading data and setting
parameters}\label{reading-data-and-setting-parameters}

The user can select to upload data as either a single set of residuals
or a pre-computed lineup of residuals and nulls. Here we illustrate use
with lineup example data sets (Figure~\ref{fig-autovi-web-setup}). To
use the lineup example data, click the ``Use Lineup Example'' button.
The data status will then update to show the number of rows and columns
in the dataset, and the CSV type will automatically be selected to the
correct option. Since the example dataset follows the variable naming
conventions assumed by the web application, the columns for fitted
values, residuals, and labels of residual plots are set automatically
(middle image). If the user is working with a custom dataset, these
options must be set accordingly. Regardless of the dataset, the user
must manually select the label for the true residual plot, whenever a
lineup file is provided, as the web application allows assessments
without this label. The last step is to click the play button (right
image) to start the assessment.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-setup-1.png}

@@

\caption{\label{fig-autovi-web-setup}To begin the workflow for
\texttt{autovi} using the lineup example dataset, the user clicks the
``Use Lineup Example'' button (left) to load the example dataset, during
which the data status and CSV type will be automatically updated. The
target columns are selected automatically, though the user must manually
select the label for the true residual plot (middle), as the web
application permits assessment without this label. The user initiates
the assessment of the lineup example data by clicking the run button
(right).}

@@

\subsubsection{Results provided}\label{results-provided}

Results are provided in multiple panels. The first row of the table
Figure~\ref{fig-autovi-web-lineup} is the most crucial to check, as it
provides the VSS and the rank of the true residual plot among the other
plots. The summary text beneath the table provides the \(p\)-value,
which can be used for quick decision-making. The lineup is for manual
inspection, and the user should see if the true residual plot is
visually distinguishable from the other plots, to confirm if the model
violation is serious.

The density plot in Figure~\ref{fig-autovi-web-distributions} offers a
more robust result, allowing the user to compare the distribution of
bootstrapped VSS with the distribution of null VSS. Finally, the
grayscale attention map (right image) can be used to check if the target
visual features, like the non-linearity present in the lineup example,
are captured by the computer vision model, ensuring the quality of the
assessment.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-lineup-1.png}

@@

\caption{\label{fig-autovi-web-lineup}Results for the lineup. The VSS of
the true residual plot is displayed in the first row of the table of VSS
values for all the null plots (left image), with a summary text beneath
the table providing the \(p\)-value to aid in decision-making. A lineup
of residual plots allows for manual inspection (right image).}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-distributions-1.png}

@@

\caption{\label{fig-autovi-web-distributions}Summaries assessing the
strength of the pattern and which elements of the plot contribute. The
density plot helps verify if the bootstrapped distribution differs from
the null distribution (left image). The attention map (right image)
offers insights into whether the computer vision model has captured the
intended visual features of the true residual plot.}
